---
title: "Poll Predictions and Errors"
author: "Shiro Kuriwaki\n"
date: "README updated: `r format(Sys.time(), '%Y-%m-%d')`"
output: 
    github_document:
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
```


# Output

The final dataset (`pres16_state.csv`) is a spreadsheet of the 50 states and DC. 


```{r, message=FALSE}
read_csv("data/output/pres16_state.csv")
```


The main columns are 

Identifiers

* `state`: Name of state (full name)
* `st`: Name of state (abbreviation)
* `color`: Outcome of the Race as defined by Cook Political: `R` (Republican), `D` (Democrat), `swing` (swing -- see details below.)

Outcomes (including estimates of VAP/VEP)

* `vap`: Estimated Voting Age Population (see below for definition)
* `vep`: Estimated Voting Eligible Population (see below for definition)
* `votes_hrc`: Votes for Clinton
* `votes_djt`: Votes for Trump
* `tot_votes`: Ballots cast for the Office of President.
* `pct_hrc_voters`: Election Outcome. Hillary Clinton's Vote as a Percentage of Ballots Cast for President. Computed by `votes_hrc / tot_votes`
* `pct_djt_voters`: Same but with Trump
* `pct_hrc_vep`: Hillary Clinton's Vote as a Percentage of Ballots (estimated) eligible population. Computed by `votes_hrc / vep`
* `pct_djt_vep`: Same but with Trump

Poll estimates. Construction detailed below and in `03_tabulate_polls.R`

Denominator of sample mean

* `cces_n_voters`: CCES sample size adjusted for estimated turnout propensity
* `cces_n_raw`: CCES raw number of respondents, or unadjusted proxy estimate of eligible population
* `cces_n_vv`: CCES number of respondents who matched to the 2016 voter file. This serves as the "true" voters (those who turned out) within the sample. 
* `cces_n_post_voters`: CCES number of respondents who responded to the post-election survey _and_ reported having voted in the 2016 presidential.

Numerator of sample mean

* `cces_tothrc_adj_trn`: CCES estimated Clinton votes 
* `cces_tothrc_raw`: CCES unadjusted total Clinton votes
* `cces_tothrc_vv`: CCES total Clinton support among validated voters.

Sample mean estimates

* `cces_pct_hrc_voters`: CCES estimated percent of Clinton votes among voters adjusting for turnout (`cces_tothrc_adj_trn/ cces_n_voters`) 
* `cces_pct_hrc_vep`: CCES estimated percent of Clinton votes among voting eligible population (`cces_tothrc_adj_trn / cces_n_raw`) 
* `cces_pct_hrc_raw`: CCES estimated percent of Clinton votes without any adjustment (`cces_tothrc_raw/ cces_n_raw`)
* `cces_pct_hrc_vv`: CCES estimated percent of Clinton votes among validated voters (`cces_tothrc_vv/ cces_n_vv`)
* `cces_pct_hrc_post_voters`: CCES estimated percent of Clinton votes among those who reported voting in the post-election survey wave (`cces_tothrc_raw_post/ cces_n_post_voters`)
* `cces_*djt*`: All same as above but with Trump
* `cces_*hrcdund*`: All same as above but with coding Undecided Democrats Clinton supporters as well.
* `cces_*djtrund*`: All same as above but with coding Undecided Republicans as Trump supporters as well.
* `cces_*hrcund*`: All same as above but with coding all Undecideds as Clinton supporters as well.
* `cces_*djtund*`: All same as above but with coding all Undecideds as Trump supporters as well.

* `cv_turnout_wgt` is the coefficient of variation on weights for each state. It is a statistic from `turnout_wgt` in `data/input/cces2016_slim.Rds`. This is used as an adjustment when calculating our parameter estimate. 


Other poll 

* `yougov_pct_hrc`: YouGov estimated of Clinton votes among voters
* `yougov_pct_djt`: Same but with Trump
* `yougov_n`: YouGov poll sample size


Parameter Estimates

* `rho_hrc_vot`: The $\rho$ parameter with Clinton support as the quantity of interest and _voters_ as the target population
* `rho_hrc_vep`: The $\rho$ parameter with Clinton support as the quantity of interest and _eligible population_ as the target population
* `rho_hrc_vvt`: The $\rho$ parameter with Clinton support as the quantity of interest and _voters_ as the target population, using validated voters only.
* `rho_hrc_pst`: The $\rho$ parameter with post-election Clinton support as the quantity of interest and _voters_ as the target population.
* `rho_*djt*`: The same $\rho$ as above but for Trump.
* `cces_*hrcdund*`: The same $\rho$ as above but coding Undecided Democrats Clinton supporters as well.
* `cces_*djtrund*`: The same $\rho$ as above but coding Undecided Republicans as Trump supporters as well.
* `cces_*hrcund*`:  The same $\rho$ as above but coding all Undecideds as Clinton supporters as well.
* `cces_*djtund*`:  The same $\rho$ as above but coding all Undecideds as Trump supporters as well.

* `neff_hrc_*` is the estimated effective sample size for the appropriate sample size $n$ using the specified estimator. For example, `neff_hrc_vot` is the effective sample size of `cces_n_voters` when we estimate `rho_hrc_vot`
* `effratio_hrc_*` is the ratio of the estimated effective sample size (`neff`) over the original sample size. For example, `effratio_hrc_vot` is given by `neff_hrc_vot/cces_n_voters`
* `neff_djt_*` and `effratio_djt*` do the same, but for Trump.




In addition to the state-level data, the following tables have summary statistics and counts not included in the state-level dataset.

* `pres16_US.csv` has total counts of some of the population turnout and voting data in the U.S. population.
*  `rho_sum_stats.csv` has summary statistics of the $\rho$ estimates. Each column is an estimate of a type of rho for a particular type of state (or all states). Each row is a summary statistic, such as 10 percent, 90 percent quantiles, and means. 


# Figures 

Figures are named according to the following convention, in the following folders:

* In `figures/summ`: Summary figures
* In `figures/rho-N`: `rho-*` is a scatter plot of a `rho` estimate on the y-axis and the target $N$ on the x-axis. We take logs, and the OLS slope coefficient is printed in the bottom-right.
    + The first part of the filename indicates the estimated $\rho$ of interest and takes a `rho`-`[candidate]`-`[specification]` format. For example, `rho-hrc-vot` is the parameter for responding to vote for Hilary Clinton, where the target population is the actual voting population. `hrc` and `djt` stand for Clinton and Trump, respectively. `hcu` and `dtu` stand for Cliton + Undecideds and Trump + Undecideds, respectively.   
    + The second part of the filename indicates the subset of observations we look at. `states-D`, `states-R`, `states-swing` indicate that we limit our scatter and coefficients to Blue, Red, and swing states, respectively. `states-pos` and `states-neg` indicate that we limit our scatter and coefficients to states that have positive or negative rho, respectively. This is an important distinction because we take the absolute value of `rho` before taking the log. `states-all` uses all states. 
* In `figures/hist`: `hist_*` is a histogram of state-level parameters. Mostly these are `rho`
* In `figures/scatter`: `scatter_*` is a figure that has state-level estimates of a quantity ($\widehat{\mu}$) on the x-axis and the observed quantity ($\mu$) on the y-axis. 
* In `figures/map`: `map_*` shows a quantity of interest in a simplified map (cartogram)
* In `figures/bars`: `bars_*` shows a quantity of interest in a barplot.



# Population Data 

The data comes from three sources and is built in `01_read_data.R`

First, the population data. The U.S. does not have an official census of citizens or voting _eligible_ citizens. Numbers on voter registrants are also out-of-date in some states. Thus the denominator of interest is fairly tricky to compute.

Here we rely on Michael McDonald's estimates at <http://www.electproject.org/>

__Voting Age Population (VAP)__ is [defined](http://www.electproject.org/home/voter-turnout/faq/denominator) as follows:

> The voting-age population, known by the acronym VAP, is defined by the Bureau of the Census as everyone residing in the United States, age 18 and older. Before 1971, the voting-age population was age 21 and older for most states.

__Voting Eligible Population (VEP)__ is [defined](http://www.electproject.org/home/voter-turnout/faq/denominator) as follows: 

> The voting-eligible population or VEP is a phrase I coined to describe the population that is eligible to vote. Counted among the voting-age population are persons who are ineligible to vote, such as non-citizens, felons (depending on state law), and mentally incapacitated persons. Not counted are persons in the military or civilians living overseas.

I pulled the VAP and VEP numbers from his spreadsheet [here](https://docs.google.com/spreadsheets/d/1VAcF0eJ06y_8T4o2gvIL4YcyQy8pxb1zYkgXF76Uu1s/edit#gid=2030096602).


Next, the observed election outcome. Vote counts are reported from official election reports and measured exactly. Any final count will do; I used the Cook Political Report's spreadsheet [here](https://docs.google.com/spreadsheets/d/133Eb4qQmOxNvtesw2hdVns073R68EZx4SfCnP4IGQf8).

The column `votes_hrc` refers to the column `Clinton (D)` in the above-linked spreadsheet. `tot_votes` refers to the sum of the three columns `Clinton (D)`, `Trump (R)`, and `Others`.


`color` is taken from the Color of this Cook Political's table and their classification of Swing. "Swing state" is defined as states in which:

> "Swing State" defined as state that flipped from '12 or was decided by less than 5%.

`R` and `D` are non-swing states defined by the presidential winner. 



# Poll Data and Methods
The main poll we use is the CCES, which is one of the few polls that has raw data available in accessible form.

The  Cooperative Congressional Election Study (CCES) is one of the largest pre-election studies conducted in the 2016 election. The CCES is conducted online for the several weeks before the election.

The target population is the national adult population. Sampling is continuously adjusted in a two-step process to obtain a representative sample. "Sample matching",  Multi-level models and other weighting schemes contribute to the retained sample and computed weights. Quoting their guide:

> Sample matching is a methodology for selection of representative samples from non-randomly selected pools of respondents. It is ideally suited for Web access panels, but could also be used for other types of surveys, such as phone surveys. Sample matching starts with an enumeration of the target population. For general population studies, the target population is all adults, and can be enumerated through the use of the decennial Census or a high quality survey, such as the American Community Survey. 

> Sample selection using the matching methodology is a two-stage process. First, a random sample is drawn from the target population. We call this sample the target sample. Details on how the target sample is drawn are provided below, but the essential idea is that this sample is a true probability sample and thus representative of the frame from which it was drawn.

> Second, for each member of the target sample, we select one or more matching members from our pool of opt-in respondents. This is called the matched sample. Matching is accomplished using a large set of variables that are available in consumer and voter databases for both the target population and the opt-in panel.


Using this methodology, YouGov created a target population from past ACS data. Then, they recruited online samples:

> The sample drawn for the CCES were chosen from the YouGov Panel, along with the MyPoints, Research Now, and SSI panels using a five-way cross- classification (age x gender x race x education x state). All respondents who completed the pre-election survey were re-invited to the post-election survey. The final set of completed pre-election interviews (numbering approximately 87,389, after quality controls were applied) were then matched to the target frame, using a weighted Euclidean distances metric.


The final (after sample matching) number of respondents in the [dataverse release](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi%3A10.7910/DVN/GDF6Z0) is 64,600. This differs from the 117,316 or the 84,292 numbers in the [pre-election press release](https://cces.gov.harvard.edu/news/cces-pre-election-survey-2016) because, among other reasons, in the interim the sample was culled of incomplete or bad-quality responses. There may have been adjustments to the case selection in order to achieve the qualities quoted above. 

For the purposes of this project, I estimated state-level predictions without adjusting for sampling or any other covariate adjustment weights. 

## Turnout adjustment

I did estimate voters and adjust for estimated turnout  in  variables `pct_hrc_voters` and `cces_n_voters`. These give a rough proxy for turnout by the coarse weights from the vote intent question. The weights are as specified in the CCES press release: 

> [A weight of] 1 for those who have already voted, 0.9 for those "definitely voting", 0.3 for those "probably voting", and 0.1 for those not sure). 

I coded other responses (No, Skipped, Missing).. to 0.


The code is in `03_tabulate_polls.R`. For simplicity I only show the estimates for the sample size and Clinton; estimates for Trump are symmetrical. 

## Validated Vote

The survey vendor, such as YouGov, has information on the survey respondents' name, geography, and demographics. This allows matching between vote files: a list of registered voters and their turnout history that states must make public. Voter files by definition show individual-level turnout for each voter and for each election. Who one votes for is secret, but whether or not one has voted is public information. While the CCES and users of their data do not know the names of the respondents, a data-vendor conducts a properietary match betch between the CCES and the voter file database to append a column of "general election vote hisotry". This allows us to estimate support _among_ respondents who was matched to a voter file and was recorded as having voted in  the 2016 general election.

## Post-election Survey

The CCES runs a second-wave, follow up survey to _respondents of the pre-election_ survey. This follow up survey includes additional question as well as questions about one's voting experience and reported vote. This allows us to estimate support who respondent to the post-election survey and record their response to the question, 'who did you vote for', rather than 'who would you support for?'


## YouGov estimates

Also, YouGov runs the CCES survey and generates their estimates with their own algorithm (not public). I took estimates from their November 16, 2016 press release [here](https://cces.gov.harvard.edu/news/cces-pre-election-survey-2016). A Google Sheets version of the same table in the release is [here](https://docs.google.com/spreadsheets/d/1pJLEHfvCN-eX1mBfe6sgs0dwF2oq9G1FcUhKFk0Pe8g).


Read the press release and guides (e.g. for 2014: [dataverse](https://dataverse.harvard.edu/file.xhtml?fileId=2794577&version=RELEASED&version=.0)) for more details on implementation.



# Estimate Performance
How do our simple state-level estimates from the CCES compare to some actual election returns?

```{r, message=FALSE, echo = FALSE}
library(ggplot2)
library(ggrepel)
library(scales)
library(readr)

df <- read_csv("data/output/pres16_state.csv")

colorvec <- c("R" =  '#d7191c', "swing" = 'darkgreen', "D" = '#2c7bb6')

gg0 <- ggplot(df, aes(x = cces_pct_hrc_voters, y = pct_hrc_voters, color = color, size = vap)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1), label = percent) +
  scale_y_continuous(limits = c(0, 1), label = percent) +
  scale_color_manual(values = colorvec) +
  guides(size = FALSE, color = FALSE) +
  coord_equal() +
  theme_bw() +
  geom_point(alpha = 0.8) +
  theme(plot.caption = element_text(size = 8)) +
  labs(caption = "Source: CCES.\nSized proportional to population.\n States colored by R (red) or D (blue) or swing (green).")


gg_hrc <- gg0 +
  annotate("text", x = 0.8, y = 0.1, label = "Poll overestimated\nClinton support", color = "darkgray") +
  annotate("text", x = 0.2, y = 0.9, label = "Poll underestimated\nClinton support", color = "darkgray") +
  labs(y = "Final Clinton Popular Vote Share")
gg_djt <- gg0 + aes(y = pct_djt_voters) +
  annotate("text", x = 0.8, y = 0.1, label = "Poll overestimated\nTrump support", color = "darkgray") +
  annotate("text", x = 0.2, y = 0.9, label = "Poll underestimated\nTrump support", color = "darkgray") +
  labs(y = "Final Trump Popular Vote Share")

```


## Estimating Clinton Support
```{r}
gg_hrc + aes(x = cces_pct_hrc_voters) +
  xlab("Turnout-adjusted Poll Estimate, Clinton Support")
```

```{r}
gg_hrc + aes(x = cces_pct_hrc_raw) +
  xlab("Raw Poll Estimate, Clinton Suport")
```

```{r}
gg_hrc + aes(x = cces_pct_hrc_vv) +
  xlab("Poll Estimate among Validated Voters, Clinton Suport")
```


## Estimating Trump Support
```{r}
gg_djt + aes(x = cces_pct_djt_voters) +
  xlab("Turnout-adjusted Poll Estimate, Trump Support")
```

```{r}
gg_djt + aes(x = cces_pct_djt_raw) +
  xlab("Raw Poll Estimate, Trump Suport")

```

```{r}
gg_djt + aes(x = cces_pct_djt_vv) +
  xlab("Poll Estimate among Validated Voters, Trump Suport")
```


# Estimates of $\rho$

We use this function for the columns `rho_*`. Input is defined in terms of its sampling target (actual votes vs. vep). We use the CCES estimates for now.
```{r, eval = FALSE}
rho_estimate <- function(data = df, N, mu, muhat, n, cv = NULL) {
  
  
  N <- data[[N]]
  n <- data[[n]]
  mu <- data[[mu]]
  muhat <- data[[muhat]]
  if (!is.null(cv)) cv <- data[[cv]]
  
  ## parts
  one_over_sqrtN <- 1 / sqrt(N)
  diff_mu <- muhat - mu
  f <- n / N
  one_minus_f <- 1 - f
  s2hat <- mu * (1 - mu)
  if (!is.null(cv)) {
    A <- sqrt(1 + (cv^2 / one_minus_f))
    one_over_A <- 1 /A
  }

  
  ## estimate of rho
  if (!is.null(cv))
    return(one_over_A* one_over_sqrtN * diff_mu / sqrt((one_minus_f / n) * s2hat))
  
  if (is.null(cv))
     return(one_over_sqrtN * diff_mu / sqrt((one_minus_f / n) * s2hat))
}
```

We present four versions of this estimate based on the sampling.

## Based on voters
For $\rho_{\text{HRC, voter}}$, we use the turnout-adjusted estimate of the candidates voteshare. Because we use turnout-adjustment, we insert the coefficient of variation of turnout weights. 

```{r, eval = FALSE}
df$rho_vot <- rho_estimate(N = "tot_votes",
                           mu = "pct_hrc_voters",
                           muhat = "cces_pct_hrc_voters",
                           n = "cces_n_voters",
                           cv = "cv_turnout_wgt")
```

## Based on voting eligible population
For $\rho_{\text{HRC, vep}}$, we use the raw estimate of the candidate's voteshare. We use the raw in order to keep the denominator as our estimate of the voter eligible population. $n$ becomes the simple number of respondents.

```{r, eval = FALSE}
df$rho_vep <- rho_estimate(N = "vep",
                           mu = "pct_hrc_voters",
                           muhat = "cces_pct_hrc_raw",
                           n = "cces_n_raw")
```

## Based on data from validated voters
For $\rho_{\text{HRC, vvt}}$ we use the estimates from the validated vote subset of the pre-election survey. Because turnout here is virtually certain, we do not adjut by turnout weights. $n$ becomes the simple number of validated voters. 

```{r, eval = FALSE}
df$rho_vvt <- rho_estimate(N = "tot_votes",
                           mu = "pct_hrc_voters",
                           muhat = "cces_pct_hrc_vv",
                           n = "cces_n_vv")
```

## Based on data from post-election survey
For $\rho_{\text{HRC, post}}$ we use the estimates from those who respondend to the post-election follow-up wave. We only use those who reported having voted, so there is no turnout-adjustment via weights here as well. $n$ becomes the number of respondents in the post-election wave who self-reported turning out to vote. 

```{r, eval = FALSE}
rho_hrc_pst = rho_estimate(N = "tot_votes",
                           mu = "pct_hrc_voters",
                           muhat = "cces_pct_hrc_voters_post",
                           n = "cces_n_post_voters")
```



I did the same for Trump voters ( $\rho_{\text{DJT, voter}}$,  $\rho_{\text{DJT, voter}}$, $\rho_{\text{DJT, vvt}}$, $\rho_{\text{DJT, post}}$), where all estimates of Clinton were replaced with their Trump equivalents. I also did the same by classifying "Undecideds" in the pre-election poll as either Clinton or Trump (the only exception here is that we do not repeat this for the post-election wave -- there were no undecideds there by definition).


# References

* McDonald, Michael P. 2017. "2016 November General Election" _United States Elections Project._ Accessed July 23, 2017.<http://www.electproject.org/home/voter-turnout/voter-turnout-data>
* CCES. 2016. Press Release. <https://cces.gov.harvard.edu/news/cces-pre-election-survey-2016>
* CCES. 2016. Data Release. <https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi%3A10.7910/DVN/GDF6Z0>
* Cook Political Report. 2017. <http://cookpolitical.com/story/10174>

