---
title: "Poll Predictions and Errors"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
```


This data combines three state-level datasets on the 2016 Presidential Election.

# Output

The final dataset (`pres16_state.csv`) is a spreadsheet of the 50 states and DC. 


```{r, message=FALSE}
read_csv("data/output/pres16_state.csv")
```


The main columns are 

Identifiers

* `state`: Name of state (full name)
* `st`: Name of state (abbreviation)
* `color`: Outcome of the Race as defined by Cook Political: `R` (Republican), `D` (Democrat), `swing` (swing -- see details below.)

Outcomes (including estimates of VAP/VEP)

* `vap`: Estimated Voting Age Population (see below for definition)
* `vep`: Estimated Voting Eligible Population (see below for definition)
* `votes_hrc`: Votes for Clinton
* `tot_votes`: Ballots cast for the Office of President.
* `pct_hrc_voters`: Election Outcome. Hillary Clinton's Vote as a Percentage of Ballots Cast for President. Computed by `votes_hrc / tot_votes`
* `pct_hrc_vep`: Hillary Clinton's Vote as a Percentage of Ballots (estimated) eligible population. Computed by `votes_hrc / vep`

Poll estimates. Construction detailed below and in `03_tabulate_polls.R`

* `cces_n_voters`: CCES sample size adjusted for estimated turnout propensity
* `cces_n_raw`: CCES raw number of respondents, or unadjusted proxy estimate of eligible population
* `cces_tothrc_adj_trn`: CCES estimated Clinton votes 
* `cces_tothrc_raw`: CCES unadjusted total Clinton votes
* `cces_pct_hrc_voters`: CCES estimated percent of Clinton votes among voters adjusting for turnout (`cces_tothrc_adj_trn/ cces_n_voters`) 
* `cces_pct_hrc_vep`: CCES estimated percent of Clinton votes among voting eligible population (`cces_tothrc_adj_trn / cces_n_raw`) 
* `cces_pct_hrc_raw`: CCES estimated percent of Clinton votes without any adjustment (`cces_tothrc_raw/ cces_n_raw`)
* `cv_turnout_wgt` is the coefficient of variation on weights for each state. It is a statistic from `turnout_wgt` in `data/input/cces2016_slim.Rds`. This is used as an adjustment when calculating our parameter estimate. 


* `yougov_pct_hrc`: YouGov estimated of Clinton votes among voters
* `yougov_n`: YouGov poll sample size


Parameter Estimates

* `rho_voter`: The $\rho$ parameter with _voters_ as the target population
* `rho_vep`: The $\rho$ parameter with _eligible population_ as the target population


# Data Sources 

The data comes from three sources and is built in `01_read_data.R`

## Observed Values

First, the sample space. The U.S. does not have an official census of citizens or voting _eligible_ citizens. Numbers on voter registrants are also out-of-date in some states. Thus the denominator of interest is fairly tricky to compute.

Here we rely on Michael McDonald's estimates at <http://www.electproject.org/>

__Voting Age Population (VAP)__ is [defined](http://www.electproject.org/home/voter-turnout/faq/denominator) as follows:

> The voting-age population, known by the acronym VAP, is defined by the Bureau of the Census as everyone residing in the United States, age 18 and older. Before 1971, the voting-age population was age 21 and older for most states.

__Voting Eligible Population (VEP)__ is [defined](http://www.electproject.org/home/voter-turnout/faq/denominator) as follows: 

> The voting-eligible population or VEP is a phrase I coined to describe the population that is eligible to vote. Counted among the voting-age population are persons who are ineligible to vote, such as non-citizens, felons (depending on state law), and mentally incapacitated persons. Not counted are persons in the military or civilians living overseas.

I pulled the VAP and VEP numbers from his spreadsheet [here](https://docs.google.com/spreadsheets/d/1VAcF0eJ06y_8T4o2gvIL4YcyQy8pxb1zYkgXF76Uu1s/edit#gid=2030096602).


Next, the observed election outcome. Vote counts are reported from official election reports and measured exactly. Any final count will do; I used the Cook Political Report's spreadsheet [here](https://docs.google.com/spreadsheets/d/133Eb4qQmOxNvtesw2hdVns073R68EZx4SfCnP4IGQf8).

The column `votes_hrc` refers to the column `Clinton (D)` in the above-linked spreadsheet. `tot_votes` refers to the sum of the three columns `Clinton (D)`, `Trump (R)`, and `Others`.


`color` is taken from the Color of this Cook Political's table and their classification of Swing. "Swing state" is defined as states in which:

> "Swing State" defined as state that flipped from '12 or was decided by less than 5%.

`R` and `D` are non-swing states defined by the presidential winner. 



# Poll Prediction

__CCES__

The  Cooperative Congressional Election Study (CCES) is one of the largest pre-election studies conducted in the 2016 election. The CCES is conducted online for the several weeks before the election.

The target population is registered voters. Sampling is continuously adjusted to obtain a representative sample. Multi-level models and other weighting schemes contribute to final state-level estimates. 

I estimated state-level predictions without adjusting for sampling or any other covariate adjustment weights. 

I did estimate voters and adjust for estimated turnout  in  variables `pct_hrc_voters` and `cces_n_voters`. These give a rough proxy for turnout by the coarse weights from the vote intent question. The weights are as specified in the CCES press release: 

> [A weight of] 1 for those who have already voted, 0.9 for those "definitely voting", 0.3 for those "probably voting", and 0.1 for those not sure). 

I coded other responses (No, Skipped, Missing).. to 0.


The code is in `03_tabulate_polls.R`,
```{r, eval = FALSE}
tab_cc <- cc_raw %>% 
  group_by(state) %>% 
  summarize(cces_n_raw = n(),
            cces_n_voters = sum(turnout_wgt, na.rm = TRUE),
            cces_tothrc_raw = sum(vote_hrc, na.rm = TRUE),
            cces_tothrc_adj_trn = sum(vote_hrc*turnout_wgt, na.rm = TRUE),
            sd_turnout_wgt = sqrt(sum((turnout_wgt - mean(turnout_wgt))^2)/n()),
            cv_turnout_wgt = sd_turnout_wgt / mean(turnout_wgt)) %>%
  mutate(cces_pct_hrc_raw = cces_tothrc_raw / cces_n_raw,
         cces_pct_hrc_vep = cces_tothrc_adj_trn / cces_n_raw,
         cces_pct_hrc_voters = cces_tothrc_adj_trn / cces_n_voters)

```



__YouGov Release__ 

YouGov runs the CCES survey and generates their estimates with their algorithm. I took estimates from their November 16, 2016 press release [here](https://cces.gov.harvard.edu/news/cces-pre-election-survey-2016). A Google Sheets version of the same table in the release is [here](https://docs.google.com/spreadsheets/d/1pJLEHfvCN-eX1mBfe6sgs0dwF2oq9G1FcUhKFk0Pe8g).




Read the press release and guides (e.g. for 2014: [dataverse](https://dataverse.harvard.edu/file.xhtml?fileId=2794577&version=RELEASED&version=.0)) for more details on implementation.



# Comparisons
Setup plot...
```{r, message=FALSE, echo = FALSE}
library(ggplot2)
library(ggrepel)
library(scales)
library(readr)

df <- read_csv("data/output/pres16_state.csv")

colorvec <- c("R" =  '#d7191c', "swing" = 'darkgreen', "D" = '#2c7bb6')

gg0 <- ggplot(df, aes(x = cces_pct_hrc_voters, y = pct_hrc_voters, color = color)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1), label = percent) +
  scale_y_continuous(limits = c(0, 1), label = percent) +
  scale_color_manual(values = colorvec) +
  guides(size = FALSE, color = FALSE) +
  coord_equal() +
  theme_bw() +
  geom_point(alpha = 0.8) +
  theme(plot.caption = element_text(size = 8)) +
  labs(caption = "Source: CCES.\nSized proportional to population.\n States colored by R (red) or D (blue) or swing (green).")


ggpres <- gg0 +
  annotate("text", x = 0.8, y = 0.1, label = "Poll overestimated\nClinton support", color = "darkgray") +
  annotate("text", x = 0.2, y = 0.9, label = "Poll underestimated\nClinton support", color = "darkgray") +
  labs(y = "Final Clinton Popular Vote Share")
```


As points 
```{r}
ggpres + aes(x = cces_pct_hrc_voters, size = vap) +
  xlab("Turnout-adjusted Poll Estimate, Clinton Support")
```


```{r}
ggpres + aes(x = cces_pct_hrc_raw, size = vap) +
  xlab("Raw Poll Estimate, Clinton Suport")
```



# Estimates of $\rho$

We use this function for the columns `rho_*`. Input is defined in terms of its sampling target (actual votes vs. vep). We use the CCES estimates for now.
```{r, eval = FALSE}
rho_estimate <- function(data = df, N, mu, muhat, n, cv = NULL) {
  
  
  N <- data[[N]]
  n <- data[[n]]
  mu <- data[[mu]]
  muhat <- data[[muhat]]
  if (!is.null(cv)) cv <- data[[cv]]
  
  ## parts
  one_over_sqrtN <- 1 / sqrt(N)
  diff_mu <- muhat - mu
  f <- n / N
  one_minus_f <- 1 - f
  s2hat <- mu * (1 - mu)
  if (!is.null(cv)) {
    A <- sqrt(1 + (cv^2 / one_minus_f))
    one_over_A <- 1 /A
  }

  
  ## estimate of rho
  if (!is.null(cv))
    return(one_over_A* one_over_sqrtN * diff_mu / sqrt((one_minus_f / n) * s2hat))
  
  if (is.null(cv))
     return(one_over_sqrtN * diff_mu / sqrt((one_minus_f / n) * s2hat))
}
```

We present two versions of this estimate based on the sampling

## Based on voters
For $\rho_{voter}$, we use

```{r, eval = FALSE}
df$rho_voter <- rho_estimate(N = "tot_votes",
                             mu = "pct_hrc_voters",
                             muhat = "cces_pct_hrc_voters",
                             n = "cces_n_voters",
                             cv = "cv_turnout_wgt")
```



## Based on voting eligible population
For $\rho_{vep}$, we use
```{r, eval = FALSE}
df$rho_vep <- rho_estimate(N = "vep",
                           mu = "pct_hrc_voters",
                           muhat = "cces_pct_hrc_raw",
                           n = "cces_n_raw")
```


Figures as PDFs are in `figures`.

# References

* McDonald, Michael P. 2017. "2016 November General Election" _United States Elections Project._ Accessed July 23, 2017.
* CCES. 2016. Press Release. <https://cces.gov.harvard.edu/news/cces-pre-election-survey-2016>
* Cook Political Report. 2017. <http://cookpolitical.com/story/10174>

