---
title: "Poll Predictions and Errors"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
```


This data combines three state-level datasets on the 2016 Presidential Election.

# Output

The final dataset (`pres16_state.csv`) is a spreadsheet of the 50 states and DC. 


```{r, message=FALSE}
read_csv("data/output/pres16_state.csv")
```


The main columns are 

* `state`: Name of state (full name)
* `st`: Name of state (abbreviation)
* `vap`: Estimated Voting Age Population
* `vep`: Estimated Voting Eligible Population
* `pct_hrc`: Election Oucome. Hillary Clinton's Vote as a Percentage of Ballots Cast for President
* `votes_hrc`: Votes for HRC. (Numerator for `pct_hrc`)
* `tot_votes`: Ballots cast for the Office of President (Denominator for `pct_hrc`)
* `cces_hrc`: Poll estimated percent of HRC support. See more below.
* `cces_n`: Sample size for poll. 
* Other columns: unedited and unnamed columns from data sources.




# Data Sources 

The data comes from three sources and is built in `01_readdata.R`

## Denominators (VAP, VEP, ..)

The U.S. does not have an official census of citizens or voting _eligible_ citizens. Numbers on voter registrants are also out-of-date in some states. Thus the denominator of interest is fairly tricky to compute.

Here we rely on Michael McDonald's estimates at <http://www.electproject.org/>

__Voting Age Population (VAP)__ is [defined](http://www.electproject.org/home/voter-turnout/faq/denominator) as folllows:

> The voting-age population, known by the acronym VAP, is defined by the Bureau of the Census as everyone residing in the United States, age 18 and older. Before 1971, the voting-age population was age 21 and older for most states.

__Voting Eligible Population (VEP)__ is [defined](http://www.electproject.org/home/voter-turnout/faq/denominator) as follows: 

> The voting-eligible population or VEP is a phrase I coined to describe the population that is eligible to vote. Counted among the voting-age population are persons who are ineligible to vote, such as non-citizens, felons (depending on state law), and mentally incapacitated persons. Not counted are persons in the military or civilians living overseas.

I pulled the numbers from his spreadsheet [here](https://docs.google.com/spreadsheets/d/1VAcF0eJ06y_8T4o2gvIL4YcyQy8pxb1zYkgXF76Uu1s/edit#gid=2030096602)


## Numerators (Votes, Percentages)

Vote counts are reported from official election reports and measured exactly. Any final count will do; I used the Cook Political Report's spreadsheet [here](https://docs.google.com/spreadsheets/d/133Eb4qQmOxNvtesw2hdVns073R68EZx4SfCnP4IGQf8)

The column `votes_hrc` refers to the column `Clinton (D)` in the above-linked spreadsheet. `tot_votes` refers to the sum of the three columns `Clinton (D)`, `Trump (R)`, and `Others`.


## Poll Prediction

### CCES
The  Cooperative Congressional Election Study (CCES) is one of the largest pre-election studies conducted in the 2016 election. I took estimates from their November 16, 2016 press release [here](https://cces.gov.harvard.edu/news/cces-pre-election-survey-2016). A Google Sheets version of the same table in the release is [here](https://docs.google.com/spreadsheets/d/1pJLEHfvCN-eX1mBfe6sgs0dwF2oq9G1FcUhKFk0Pe8g).

The CCES is conducted online for the several weeks before the election.

The target population is registered voters. Sampling is continuouslly adjusted to obtain a representative sample. Multi-level models and other weighting schemes contribute to final state-level estimates. 

Read the press release and guides (e.g. for 2014: [dataverse](https://dataverse.harvard.edu/file.xhtml?fileId=2794577&version=RELEASED&version=.0)) for more details on implementation.



# Comparisons
```{r, message=FALSE}
library(ggplot2)
library(ggrepel)
library(scales)
library(readr)

df <- read_csv("data/output/pres16_state.csv")

ggplot(df, aes(x = cces_hrc, y = pct_hrc, size = cces_n)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_label_repel(aes(label = st), segment.alpha = 0.5) +
  scale_x_continuous(limits = c(0, 1), label = percent) +
  scale_y_continuous(limits = c(0, 1), label = percent) +
  guides(size = FALSE) +
  coord_equal() +
  theme_bw() +
  labs(x = "CCES Pre-election Survey Clinton Support",
       y = "Final Clinton Popular Vote Share",
       caption = "Sized by survey sample size")

```





# References

* McDonald, Michael P. 2017. "2016 November General Election" _United States Elections Project._ Accessed July 23, 2017.
* CCES. 2016. Press Release. <https://cces.gov.harvard.edu/news/cces-pre-election-survey-2016>
* Cook Political Report. 2017. <http://cookpolitical.com/story/10174>

